Actuator line modeling of vertical-axis turbines

### 2016-08-27 reviews/reviewer2-response.md

# Response to reviewer 2

>1. Pag1. Please insert the list of Symbols since there are a lot of
abbreviations in the text, a quick reference could be useful.

>2. Pag. Pag3. Sect. 2. The parameter reduced frequency "k" is the key parameter
for unsteadiness considerations concerning VAT: some more detailed comments
should be given concerning the actual value of this parameter in the case study
numerically modelled. At the same time a comment should be given concerning the
ALM capabilities (and/or future improvements) in terms of unsteady aerodynamics.

>3. Pag4. Sect.4.  Is the force smoothing with the Gaussian function really
necessary? In P.-E. Réthoré and N. N. Sørensen, A discrete force allocation
algorithm, Wind Energy, 2012. Rethore and Sorensen showed how this spatial force
filter could be avoided with modified Rhie–Chow interpolation algorithm.

In our experience, especially with the finer LES grids, the force smoothing is
necessary. Without smoothing, the blade loading becomes quite "noisy" due to the
subsequent instability from sharper momentum source gradients. In this study we
did not investigate changes to the numerical algorithms used to solve the flow
field, but certainly future research may prove the Gaussian width can be
narrowed or eliminated if, for example, modified Rhie--Chow interpolation is
implemented in OpenFOAM.

>4. Pag6. Sect.7. Does the authors took into consideration that the ALM could be
in theory able to catch the Tip vortex effect without a correction model? i.e.
the vortex generated by the lifting of the blade line is already influencing the
angle of attack on the blade.

>5. Pag7. Sect.8-9 . Add a sketch of the domain boundaries with the position of
the turbine, and the flow direction. Figure 3 is not very effective in terms of
quantitative information on the actual implementation of the code.

>6. Sect. 8-9 The details on the domain boundaries of the implemented case study
should be an opportunity for clarify if any blockage effect was present in the
reference experiments and if the same blockage conditions were adopted in the
numerical implementation.

>7. Pag12. Sect.9.4 "The near-wake's mean velocity field at x/D=1 is shown in
Figure 12..." should be actually Figure 13

### 2016-08-27 paper.tex
Add nomenclature per reviewer 2 request

@@

@@

@@

@@

Actuator line modeling (ALM), originally developed by Sorensen and Shen
\cite{Sorensen2002}, is an unsteady method that tracks blade element locations,
and has become popular for modeling axial-flow or horizontal-axis turbines, and
has been shown in blind tests to be competitive with blade-resolved
CFD~\cite{Krogstad2013, Pierella2014}. The ALM combined with large eddy
simulation has become the state-of-the-art for modeling entire wind
farms~\cite{Archer2013, Churchfield2012, Sorensen2015, Fleming2013,
Fleming2014}. Like other blade element techniques, the effectiveness of the ALM
for AFTs is in part due to the quasi-steady nature of the flow in the blade
reference frame, and the relatively rare occurrence of stall. Note that a
similar method can be used with the Navier--Stokes equations in
vorticity--velocity form~\cite{Scheurich2011b}, which may be more efficient when
vorticity is confined to a relatively small fraction of the domain, e.g., for a
standalone turbine with a uniform laminar inflow.

The ASSM and ALM were implemented to model a very low Reynolds number (based on
rotor diameter, $Re_D \sim 10^4$) 2-D cross-flow turbine experiment in a flume
using large eddy simulation (LES)~\cite{Shamsoddin2014}. Performance predictions
for this case were not reported, but the ALM was shown to be more effective at
postdicting the wake characteristics measured in the experiments by Brochier et
al.~\cite{Brochier1986}.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

\usepackage{nomencl}
\makenomenclature

@@

\printnomenclature

@@

\nomenclature{CFT}{Cross-flow turbine}
\nomenclature{VAT}{Vertical-axis turbine}
\nomenclature{AFT}{Axial-flow turbine}
\nomenclature{HAT}{Horizontal-axis turbine}

@@

\nomenclature{RANS}{Reynolds-averaged Navier--Stokes}

@@

Actuator line modeling (ALM), originally developed by Sorensen and Shen
\cite{Sorensen2002}, is an unsteady method that tracks blade element locations,
and has become popular for modeling axial-flow or horizontal-axis turbines, and
has been shown in blind tests to be competitive with blade-resolved
CFD~\cite{Krogstad2013, Pierella2014}. The ALM combined with large eddy
simulation (LES) has become the state-of-the-art for modeling entire wind
farms~\cite{Archer2013, Churchfield2012, Sorensen2015, Fleming2013,
Fleming2014}. Like other blade element techniques, the effectiveness of the ALM
for AFTs is in part due to the quasi-steady nature of the flow in the blade
reference frame, and the relatively rare occurrence of stall. Note that a
similar method can be used with the Navier--Stokes equations in
vorticity--velocity form~\cite{Scheurich2011b}, which may be more efficient when
vorticity is confined to a relatively small fraction of the domain, e.g., for a
standalone turbine with a uniform laminar inflow.

\nomenclature{ALM}{Actuator line model}
\nomenclature{LES}{Large eddy simulation}

The ALM has been previously used to model a very low Reynolds number (based on
rotor diameter, $Re_D \sim 10^4$) 2-D cross-flow turbine experiment in a flume
using large eddy simulation (LES)~\cite{Shamsoddin2014}. Performance predictions
for this case were not reported, but the ALM was shown to be more effective at
postdicting the wake characteristics measured in the experiments by Brochier et
al.~\cite{Brochier1986}.

@@

\nomenclature{$c$}{Blade chord length}
\nomenclature{$R$}{Maximum rotor radius}
\nomenclature{$Re_D$}{Rotor diameter Reynolds number}
\nomenclature{DOE}{US Department of Energy}
\nomenclature{SNL}{Sandia National Laboratories}

@@

\nomenclature{$\vec{U}_\mathrm{rel}$}{Blade element local relative velocity}
\nomenclature{$\alpha$}{Blade element angle of attack}
\nomenclature{$\omega$}{Rotor angular velocity}
\nomenclature{$r$}{Blade element radius}
\nomenclature{$\vec{U}_\mathrm{in}$}{Blade element local inflow velocity}

@@

\nomenclature{$U_\infty$}{Free stream velocity}

@@

\nomenclature{$\lambda$}{Rotor tip speed ratio}
\nomenclature{$k$}{Reduced frequency or turbulence kinetic energy}

@@

\nomenclature{$C_l$}{Blade element lift coefficient}
\nomenclature{$C_d$}{Blade element drag coefficient}
\nomenclature{$C_m$}{Blade element moment coefficient}

@@

\nomenclature{$Re_c$}{Blade chord Reynolds number}

@@

\nomenclature{$\epsilon$}{Gaussion force smoothing function width}

@@

\nomenclature{$C_P$}{Rotor power coefficient}
\nomenclature{$C_D$}{Overall rotor drag (a.k.a. thrust) coefficient}
\nomenclature{$D$}{Rotor diameter}

@@

\nomenclature{$H$}{Rotor height}

@@

\nomenclature{$\lambda_0$}{Tip speed ratio of maximum power coefficient}
